---
title: "Geocode Planning Data 2010-2030 - Brighton"
author: "Adam Dennett"
published-title: 26/10/24
format: 
  html:
    code-fold: true
    code-tools: true
editor: visual
---

## New Dwelling Units outside of Designated Development Areas, Brighton and Hove, 2010-2030

Data on likely growth in housing due to planned developments is available here:

<https://www.brighton-hove.gov.uk/planning/planning-policy/strategic-housing-land-availability-assessment-shlaa-update-2023>

Data table downloaded from here:

<https://www.brighton-hove.gov.uk/planning/planning-policy/data-tables-strategic-housing-land-availability-assessment-shlaa-2023>

Within Brighton, there are Development Areas where most residential development is planned to occur, but things also occur outside of these areas and this is included in table B in the datafile - this is what I am going to look at, primarily.

<https://www.brighton-hove.gov.uk/planning/planning-policy/strategic-housing-land-availability-assessment-shlaa-update-2023#tab--appendix-1>

In the code that you can reveal with the button above, I Use the Google Maps API to geocode the addresses in the data file and a bit of manual editing to clean up any errors.

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
library(ggmap)
library(chattr)
library(here)
library(tidyverse)
library(janitor)
library(tmap)
library(sf)
library(stringi)

source("config.R")

register_google(key = api_key)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
shlaa <- read_csv(here("data", "shlaa_raw.csv")) %>% 
  clean_names() %>% 
  mutate(site_long = paste0(site, ", ", "Brighton and Hove, England"))

shlaa_da <- read_csv(here("data", "shlaa_raw_da.csv")) %>% 
  clean_names() %>% 
  mutate(site_long = paste0(site_6_plus_units, ", ", "Brighton and Hove, England"))
  
  
#shlaa_geocode <- mutate_geocode(shlaa, location = paste0(Site, ", ", "Brighton and Hove, England"), output = "latlona")

shlaa_geocode <- mutate_geocode(shlaa, location = site_long, output = "latlona")

shlaa_da_geocode <- mutate_geocode(shlaa_da, location = site_long, output = "latlona")

#write_csv(shlaa_da_geocode, here("data", "shlaa_da_geocode.csv"))

# Subset where address is not NA 
shlaa_geocode_with_address <- subset(shlaa_geocode, !is.na(address)) 
# Subset where address is NA 
#shlaa_geocode_without_address <- subset(shlaa_geocode, is.na(address))

#write_csv(shlaa_geocode_with_address, here("data", "shlaa_geocode_with_address.csv"))

#write_csv(shlaa_geocode_without_address, here("data", "shlaa_geocode_without_address.csv"))

#shlaa_geocode_without_address <- mutate_geocode(shlaa_geocode_without_address, location = site, output = "latlona")

shlaa_geocode_without_address <- read_csv(here("data", "shlaa_geocode_without_address.csv"))

shlaa_geocode_with_address <- read_csv(here("data", "shlaa_geocode_with_address.csv"))

shlaa_da_geocode <- read_csv(here("data", "shlaa_da_geocode.csv"))

#write_csv(shlaa_geocode, here("data", "shlaa_geocode.csv"))
#shlaa_geocode <- read_csv(here("data", "shlaa_geocode.csv"))


shlaa_geocode <- rbind(shlaa_geocode_with_address, shlaa_geocode_without_address)
```

Dwelling units outside of Designated Development Areas, Brighton and Hove, 2010-2030

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}


# Remove rows with missing values in the 'lon' or 'lat' columns
#shlaa_geocode_clean <- shlaa_geocode %>%
#  filter(!is.na(lon) & !is.na(lat))

#write_csv(shlaa_geocode_clean, here("data", "shlaa_geocode_clean.csv"))
shlaa_geocode_clean <- read_csv(here("data", "shlaa_geocode_clean.csv"))

# Convert dataframe to sf object
shlaa_geocode_sf <- st_as_sf(shlaa_geocode_clean, coords = c("lon", "lat"), crs = 4326) %>% 
  clean_names() 

shlaa_da_geocode_sf <- st_as_sf(shlaa_da_geocode, coords = c("lon", "lat"), crs = 4326) %>% 
  clean_names() 

shlaa_geocode_sf$planning_status <- stri_enc_toutf8(shlaa_geocode_sf$planning_status, is_unknown_8bit = FALSE, validate = FALSE)

shlaa_geocode_sf$site <- stri_enc_toutf8(shlaa_geocode_sf$site, is_unknown_8bit = FALSE, validate = FALSE)



# Create the tmap
#tm_shape(shlaa_geocode_sf) +
#  tm_dots(size = "total_supply_2010_2030", col = "blue", alpha = 0.7) +
#  tm_layout(title = "Proportional Circles of Total Supply 2010-2030")

# Create the tmap
tmap_mode("view") # If you want an interactive map
tm_shape(shlaa_geocode_sf) +
  tm_dots(size = "total_supply_2010_2030", 
          col = "planning_status", 
          alpha = 0.7, 
          legend.size.show = TRUE, 
          popup.vars = c("site", "total_supply_2010_2030", "planning_status")) 


```

Dwellings Inside Designated Development Areas, Brighton and Hove, 2010-2030

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}

tm_shape(shlaa_da_geocode_sf) +
  tm_dots(size = "total_supply_2010_2030", 
          col = "planning_status", 
          alpha = 0.7, 
          popup.vars = c("site_6_plus_units", "total_supply_2010_2030", "planning_status")) 
```

## Pupil Yield

New Housing Developments will generate additional children, but how many? The Department for Education has published guidance on estimating pupil yield from housing developments. Estimates are based on the number of bedrooms in the new housing developments and the type of housing development. The estimates are based on data from the Department for Education and the Office for National Statistics.

They even provide a handy dashboard so you can estimate the number of additional children that will be generated by new housing developments in your area. The dashboard allows you to input the number of bedrooms in the new housing developments and the type of housing development and provides an estimate of the number of additional children that will be generated.

<https://department-for-education.shinyapps.io/pupil-yields-dashboard/>

<https://assets.publishing.service.gov.uk/media/64d0f71be5491a00134b5940/Estimating_Pupil_Yield_from_Housing_Development.pdf>

Estimated Pupil Yield per bedroom in Brighton and Hove in 2021/22 is:

-   Early Years: 0.0253

-   Primary: 0.0816

-   Secondary: 0.0408

-   Special Schools: 0.0035

### Likely Pupil Yield from Developments East of Longhill

Big Rottingdean / Ovingdean Developments (not sure how many bedrooms, but guess average of 2.5 bedrooms)

Site 48 - 72 St Aubyns - 90 Site 42 - 45 Ocean Hotel - 48 Site 50 - 31

286 dwellings-ish

286 \* 2.5 = 715 bedrooms

-   Early Years: 715 \* 0.0253 = 18.1

-   Primary: 715 \* 0.0816 = 58.3

-   Secondary: 715 \* 0.0408 = 29.2

-   Special Schools: 715 \* 0.0035 = 2.5

-   Total: 108.1 additional children
